---
title: "Class 3 Exercise"
author: "Names of those in your group"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Scenario

You work as a data scientist at a company that sells widgets. The CEO and owner is extremely engaged in looking at the most recent data on sales but is not a statistician and is prone to pay too much attention to meaningless day-to-day and month-to-month fluctuations.

January 2, 8 AM:  The CEO comes into your office and expresses worry about widget sales for the most recent month, December.  She thinks sales have tanked and wants you to look into the situation further and provide a brief report by noon. Widget demand is seasonal and the business depends on strong holiday sales. She wants a brief report on her desk by noon. 

## Topics

1. EDA workflow
2. Practice data manipulation and visualization
3. Statistical inference
4. Communication


## Process

- Please gather into groups to work on this project!
- Put group members' names up above in the yaml heading under "author" (where it currently says "Names of those in your group").
- Collaborate on one document.
- When you are done, compile to HTML (or PDF), and submit through Canvas.

## EDA workflow

1. Formulate a question:  Are last month's sales (month 12 of year 5) down?

2. Read in your data.  For this exercise we will simulate a dataset.

```{r}

widget <- expand.grid(year = 1:5, month = 1:12, day = 1:30)

# The expand.grid() function creates a dataframe with unique combinations of the values
# from each variable. Here we are setting up a data.frame for 5 years of data.

head(widget)

set.seed(1126) # Use set.seed() to ensure identical datasets

# Now, simulate sales data using the uniform distribution and the normal distribution:
# runif() and rnorm().  For simplicity we will pretend each month has exactly
# 30 days.

widget %<>%
  mutate(sales = ifelse(month < 12, 
                        runif(5 * 11 * 30, min = 800, max = 1200) +
                          rnorm(5 * 11 * 30, mean = 100, sd = 100),
                        runif(5 * 11 * 30, min = 1100, max = 1300) +
                          rnorm(5 * 11 * 30, mean = 100, sd = 100)),
         sales = ifelse(month == 12 & year == 5, 
                        sales - rnorm(30, mean = 100, sd = 30),
                        sales),
         year = factor(year),
         month = factor(month)) %>% 
  arrange(year, month, day) %>%
  mutate(instance = 1:(5*12*30)) # instance is a row counter.

head(widget)

```

3. Check the packaging: dim(), nrow(), ncol()

```{r}

```

4. Inspect the dataset: str(), glimpse(), View()

```{r}

```

5. Look at the top and the bottom of your data: head(), tail()

```{r}

```

6. Summarize the data: summary(), table(), hist()

```{r}

```

7. Try the easy solution first 

    + Plot daily sales
    
    ```{r}

```

    + Plot monthly sales
    
    ```{r}

```

    + Plot yearly sales
    
    ```{r}

```
   
    + Summarize total and average sales by month, and calculate confidence intervals.

```{r}

```
    
    + Summarize total and average sales by  year, and calculate confidence intervals. For the CIs remember that SEM = s / sqrt(n) and that the interval is mean -/+ 1.96 * SEM.

```{r}

```

8. Challenge your solution

    + Plot monthy sales facetted by year

```{r}

```

    + Plot monthly sales colored by year


```{r}

```

    + Summarize total and average sales by month and year, and calculate confidence intervals.


```{r}

```

9. Follow up questions. See next section.

## Statistical inference 

Answer the CEO's question:  Is there a real drop in year 5 month 12 sales or is the difference just due to random variation? 

How would you approach this question?


```{r}

```

## Communication 

Write a paragraph summary of your descriptive and inferential findings.



