---
title: "Statistics and Predictive Analytics, Lab 2"
author: "Your name here"
output:
  html_notebook
---

Welcome to Lab 2!  

- Check Canvas for the due date.
- Make sure to write your name in the yaml header above.  
- Take care that the output type in the header remains "html_notebook." (If you accidentally change it, by clicking "Knit to HTML" for example," simply change the output back to "html_notebook" and save.  This should restore the "Preview" option for you.) 
- Before compiling, click the "Run" button in the upper left of the RStudio toolbar, and select "run all."  This will ensure that your code chunks have run and will be visible to us in the compiled HTML document that you submit for the grade.
- Click the "Preview" button on the toolbar to compile your notebook into HTML.  This resulting HTML document is what you will submit through the Lab 1 assignment.
- The HTML answer key for this notebook is available for you to check, if you want to do so.  

### Introduction

The purpose of the lab is to give you practice with the skills we have covered so far in the course, from EDA to linear regression:  

- Create exploratory plots and think about data modeling.
- Fit and interpret simple and multivariable linear models.
- Interpret effect sizes.
- Calculate error metrics and compare models.
- Calculate quantities of interest.

You will be working with the Boston housing dataset that is available in the MASS package.  This dataset records tract level statistics on housing in the Boston area in the 1970s.  The outcome variable is `medv`:  Median Home Value (expressed in $1000s).  

Our goal in this analysis, specifically, will be to find out whether air pollution, encoded in the variable `nox`, impacts home values.  Are increases in nitrogen oxide associated with lower home values? Or, to put the question the other way around, is clean air associated with more expensive homes?  Data analysis can be focused on prediction or description, also known as inference.  Our goal here is the second.


```{r message=FALSE, warning=FALSE}
library(tidyverse)

# install.packages("MASS") # install the MASS package (if
# you don't already have it)

library(MASS)
library(knitr)
library(caret)
library(arm)

data(Boston) 

?Boston # Look at the data dictionary 

b <- Boston # Rename to avoid more typing than necessary
```


A more complete data dictionary is available in the original Harrison & Rubinfeld paper that used the dataset, starting on page 96: https://www.law.berkeley.edu/files/Hedonic.PDF.


### Data exploration and modeling

The `summary()` function allows you to quickly assess the distributions of the variables, including which ones are continuous and which are binary, and whether there are any missing observations.  

```{r}
summary(b)  # Get a summary of each variable.  
```

For example, we can see that "chas" is binary: it is an indicator for whether the tract (not the home) bounds the Charles River.  Another handy tool is `glimpse()` from dplyr, which will give you a sense of the structure of the data and of the observations themselves.

```{r}
glimpse(b)   # Look at the structure of the dataset
```

We can review the univariate distributions of the predictors in the dataset using histograms like this:

```{r}

for(i in 1:length(names(b))){
  hist(b[,i], main = names(b)[i])
}
```

What are we looking for exactly?  We want to be familiar with the distributions of the predictors before we jump into regression modeling to inform choices about how to clean and encode variables:  are there outliers? nonsensical values? should a variable be encoded as a factor or left numeric? would it make sense to do any other variable transformations such as turning a continuous variable into a binary or categorical variable?

For example, the rad variable (an index of the highway accessibility) has a strange distribution.  Here is the scatterplot of rad against medv:

```{r}
ggplot(b, aes(rad, medv)) +
  geom_point() +
  labs(title = "Medv ~ rad")

table(b$rad) # Here is the distribution of rad.
```

Is rad best left as numeric or should it be treated as a factor variable or should we discretize it (< 9 and > 9)?  These are the sorts of questions we need to think about at the data exploration/data modeling phase of an analytic project.

Rad = 24 is a weird value.  Nevertheless, numeric differences in rad are meaningful, in that higher levels represent less accessibility (and lower represent greater).  Harrison and Rubinfeld entered Rad into the model as a logged variable. It looks like this would actually make the fit worse, though the scatterplot looks better.  Turning rad into a factor variable fits a separate model for each level of rad.  The concern in doing that is we may be overfitting.  I would leave rad as numeric.

**Question 1**: Calculate and report average `nox` in this dataset with 95% CIs.  

```{r}
# Write your code here

```

**Question 2**: Plot the relationship between `nox` and `medv` (i.e., make a scatterplot with a least squares line).  Make sure to title the plot and label the axes. 

```{r}
# Write your code here

```

### Fitting and interpreting models

*Question 3*:  Fit a simple linear regression with medv as the outcome variable and nox as the predictor. Report the 95% confidence interval for $\hat\beta$ (the coefficient for nox) and comment on whether its difference from 0 might be due to chance. 


```{r}
# Your code goes here


```

>Write your answer here.

**Question 4**: Fit a linear model of `medv` with all the predictors (we will call this the "full model") and interpret the coefficient for `nox`.  By "interpret" I mean you should explain what the coefficient says about the relationship between  `nox` and `medv`.  Additionally, report a 95% CI for the `nox` coefficient.  

```{r}
# Write your code here. 

```

>Write your answer here.

### Effect sizes

**Question 5**: Which of the predictors in the full model has the largest effect size? 

```{r}
# Your code goes here

```

>Write your answer here.


### Error metrics and model comparison

**Question 6**:  Report the $R^2$ and RMSE of the full model.  You can report $R^2$ from the model summary, of course, but see if you can write a function to calculate $R^2$ (refer to the formulas in the lecture slides).

```{r}
# your code goes here

```

**Question 7**: Fit a KNN model using the variables the full model.  Does the KNN model fit the data better than your linear model?  Explain your reasoning. Sometimes, for example, if a non-parametric model fits the sample data better, it can be because the parametric assumptions of a linear model are being violated. (It is also possible that the non-parametric model is overfitting the sample data.  More on that in a future class.)

```{r}
# Your code goes here

```

>Write your answer here.

### Statistical communication

**Question 8**: Our original question was:  "is clean air associated with more expensive homes?"  Write a paragraph in which you address that question, including any thoughts you have about the shortcoming of the analysis or your level of confidence in the results. 

>Write your paragraph here

In class I made the point that regression model results generally should be translated into quantities of interest so that non-statistical audiences can better understand them.  

**Question 9 (optional)**: Using the full model calculate the value of a typical home at a low level of `nox` and at a high level of `nox`.  It is up to you to define what counts as "low" or "high."  Your goal should be to complete the following sentence:  "When `nox` increases from a low level (defined by ...) to a high level (defined by ...) we observe an increase/decrease in home values of, on average, ... dollars."

```{r}
# Your code goes here
```

>Complete the sentence here.


**Question 10**:  Please score yourself on this lab using the answer key.  For each question, reflect on the difficulty of the question for you and whether you:

- Did not try (just copied the answer key or left it blank).
- Looked at the answer key to get a hint.
- Tried but got the wrong answer.
- Tried and got the right answer.

>Write your answers here.

**Next step.**  After you've finished this lab, go take Quiz 2.  You have one try at the quiz, limited to 30 minutes.  There are five multiple choice and multiple answer questions focusing on the material from this lab.

